Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-11 paper/disc20clustering.tex
wip

We study two models in this paper: the \emph{general partitioning} model, and its subproblem, the \emph{learning} model.

@@

\medskip 
In both models, we assume that communication patterns are not known to our algorithms at the beginning.
We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
We emphasize that all our algorithms have a \emph{strict} competitive ratio (without an additive term).

@@

\medskip
\noindent
\textbf{Distributed implementation.}
While we have described the algorithm globally so far, we note that it allows for an efficient distributed implementations. 
The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a broadcast of the end of the phase.
We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
A merge of two components may require finding a fresh cluster (for details see the proof of Lemma~\ref{lem:1req}).
In the following, we show how to efficiently find a fresh cluster in a distributed manner.
We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
Each cluster maintains the counter of fresh clusters in its subtree.
To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
Upon encountering a fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
Summarizing, ending the phase requires a single broadcast, and merging components requires $O(\log \ell)$ communication.

>>>>

We study two models in this paper: the \emph{general partitioning} model, and its subproblem, the \emph{learning} model.
In both models, we assume that communication patterns are not known to our algorithms at the beginning.
We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
%We emphasize that all our algorithms have a \emph{strict} competitive ratio (without an additive term).

@@

@@

\medskip
\noindent
\textbf{Distributed implementation.}
While we have described the algorithm globally so far, we note that it allows for an efficient distributed implementations. 
The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a broadcast of the end of the phase.
We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
A merge of two components may require finding a fresh cluster (for details see the proof of Lemma~\ref{lem:1req}).
In the following, we show how to efficiently find a fresh cluster in a distributed manner.
We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
Each cluster maintains the counter of fresh clusters in its subtree.
To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
Upon encountering a fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
Summarizing, ending the phase requires a single broadcast, and merging components has $O(\log \ell)$ communication complexity.
